Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-05-31 doc/paper.tex
sloppy looking table seems acceptable in 2015 proceedings

As early as Jelinek's (in)famous 1980's quote, ``{\em Every time I fire a linguist, my performance goes up},'' both linguists and linguistics have been increasingly marginalized in natural language processing. 
More recently, for example, \mynewcite{collobert 2011 footnote 8} mention the possibility of ``{\em [ideally]...to learn from letter sequences rather than words},'' illustrating an AI-motivated desire to move away from even the most fundamental of linguistic concepts.
A specific instance of this overall trend is the gradual shift of machine translation (MT) away from classical linguistic rule-based approaches \mycitep{hutchins} to contemporary machine learning approaches, which underlie statistical phrase-based systems like Moses \mycitep{moses} and more recent neural network systems \mycitep{sutskever, bengio}.  
But in the end, even the most cutting-edge machine learning MT system relies crucially on multilingual parallel corpora, typically taken from external sources like multilingual parliamentary proceedings. 
This results in what we perceive to be a disproportionate amount of effort being dedicated to the machine learning component of machine translation, with the ever-important input corpora being treated almost as an afterthought.\footnote{
    Of course, these parallel corpora represent hundreds of translator-hours of work. 
    However, these translations are generally optimized for human consumption, not machine learning, with no regard for such niceties as sentence alignment.
    This is illustrated briefly in Section \ref{case_study}.
    }

@@

As a concrete example to illustrate the need for higher-quality parallel corpora, we briefly consider the official Chinese to English baseline translation of the {\small \tt tst2010} data set\footnote{https://wit3.fbk.eu/score.php?release=2014-01} from the IWSLT shared task studied in Section \ref{experiments}.
As a very crude way of assessing translation quality, we used the official evaluation software, MultEval \mycitep{multeval} to compute scores for each translated line individually.
Upon manual examination, it was found that the lines with the worst individual TER scores \mycitep{ter}, some as high as 300\%, almost all corresponded to significant translation errors in the Chinese side of the gold corpus.

@@

One would expect such outliers to be relatively well-tolerated in training sets
But the distressing part is that the {\small \tt tst2010} data set has been used as the official development/tuning set for the past several IWSLT evaluation campaigns.
It is unclear what effect such noisy corpora would have on other systems, 
especially when used for tuning
This results in a significant TER score change that is almost as large as the score difference between the highest and lowest ranking system from the evaluation campaign.  
BLEU scores, on the other hand, did not change much.

Removing TED talk number 767 (over 200 lines), as well as 15 hand-picked lines with bad TER scores,

% TODO: table and reference

@@

@@

some neural network MT paper?

aha, yoshua bengio's "curse of dimensionality" - becomes a blessing?  
Unlike many machine learning tasks (image recognition, finance), NLP data are purely artificial/manmade.  New paradigm?  Not enough attention has been paid IMO to semi-automated corpus generation.  Well, typical approaches are paid, like Mechanical Turk... but there is a vast untapped source of data in the form of student problem sets, in the spirit of reCAPTCHA (citation!?)
NLP is unique among machine learning tasks in that humans are not only consumers, but also producers of data - the alpha and the omega, the beginning and the end.
MT is further unique among NLP in that both input and output representations are widely accepted (modulo dialectical variations) - the source and target languages themselves.
People on the street would agree to the representations.

@@

>>>>

As early as Jelinek's (in)famous quote from the 1980's that ``{\em Every time I fire a linguist, my performance goes up},'' both linguists and linguistics have been increasingly marginalized in natural language processing. 
More recently, for example, \mynewcite{collobert 2011 footnote 8} mention the possibility of ``{\em [ideally]...to learn from letter sequences rather than words},'' illustrating an AI-motivated desire to move away from even the most fundamental of linguistic concepts.
A specific instance of this overall trend is the gradual shift of machine translation (MT) away from classical linguistic rule-based approaches \mycitep{hutchins} to contemporary machine learning approaches, which underlie statistical phrase-based systems like Moses \mycitep{moses} and more recent neural network systems \mycitep{sutskever, bengio}.  
But in the end, even the most cutting-edge machine learning MT system relies crucially on multilingual parallel corpora, typically taken from external sources like multilingual parliamentary proceedings. 
This results in what we perceive to be a disproportionate amount of effort being dedicated to the machine learning component of machine translation, with the ever-important input corpora being treated almost as an afterthought.\footnote{
    Of course, these parallel corpora represent hundreds of translator-hours of work. 
    However, these translations are generally optimized for human consumption, not machine learning, with no regard for such niceties as sentence alignment.
    This is illustrated briefly in Section \ref{case_study}.
    }

@@

% TODO: don't need to go into this much detail, do you? This would be the first section to trim, if it runs OVER length
% To find an example of suboptimally translated parallel corpora in the literature, one need not look far.
As a concrete example to illustrate the need for higher-quality parallel corpora, we briefly consider the official Chinese to English baseline translation of the {\small \tt tst2010} data set\footnote{https://wit3.fbk.eu/score.php?release=2014-01} from the IWSLT shared task studied in Section \ref{experiments}.
As a very crude way of assessing translation quality, we used the official evaluation software, MultEval \mycitep{multeval} to compute scores for each translated line individually.
Upon manual examination, it was found that the lines with the worst individual TER scores \mycitep{ter}, some as high as 300\%, almost all corresponded to significant translation errors in the Chinese side of the gold corpus.

@@

% TODO: table with change; also list best and worst 
% caption should mention that the difference between 
    % ALTHOUGH you would expect that it would improve all models, it's not clear whether 
Removing TED talk number 767 (over 200 lines), as well as 15 hand-picked lines with bad TER scores,

First, note that the scores improve, which is consistent with the observation that

good machine translations of bad Chinese sentences.

This results in a significant TER score change that is almost as large as the score difference between the highest and lowest ranking system from the evaluation campaign.  
BLEU scores \mycitep{bleu}, on the other hand, did not change much.

One would expect such errors to be relatively well-tolerated in large training corpora
But one bad 200-line talk constitutes a significant fraction of {\small \tt tst2010}, which is only 1500 lines total.
More distressing is the fact is that the {\small \tt tst2010} data set has been used as the official development/tuning set for the past several IWSLT evaluation campaigns.
It is unclear what effect such noisy corpora would have on other systems, 
especially when used for tuning

% TODO: table and reference
\begin{table}[]
\centering
\label{table:tst2010}
\begin{tabular}{|l|l|l|}
\hline
 & BLEU & TER  \\ \hline
Full {\small \tt tst2010} & 11.2 & 77.0  \\ \hline
Purged {\small \tt tst2010} & 11.8 & 71.3  \\ \hline
%\hline
%{\small \tt tst2014} &  &   \\ \hline
% &  &   \\ \hline
\end{tabular}
\caption{By comparison, the difference between the baseline TER score and (not shown) } % TODO: caption
\end{table}

@@

Unlike many machine learning tasks (image recognition, finance), NLP data are purely artificial/manmade.  New paradigm?  Not enough attention has been paid IMO to semi-automated corpus generation.  Well, typical approaches are paid, like Mechanical Turk... but there is a vast untapped source of data in the form of student problem sets, in the spirit of reCAPTCHA (citation!?)
NLP is unique among machine learning tasks in that humans are not only consumers, but also producers of data - the alpha and the omega, the beginning and the end.
MT is further unique among NLP in that both input and output representations are widely accepted (modulo dialectical variations) - the source and target languages themselves.
People on the street would agree to the representations.

@@

@@

training curves (just to prove my controls aren't wrong)
my synthetic corpora don't help. 
but in-domain corpus helps significantly more than out-of-domain. (this is a commonly reported result)

Bottom line
My corpora don't really seem to help, but neither do out-of-domain corpora.

this will easily take up a whole column, because it includes 2 tables...
